---
title: "Predictive model for FIFA 2019"
output: 
  html_document:
    toc: true
---

# Read data

Read the data from the kaggle website https://www.kaggle.com/karangadiya/fifa19.

```{r warning=FALSE, message=FALSE}
library("readr")
fifa19 <- as.data.frame(read_csv("fifa2019.csv"))
```

# Data Preprocessing

Transform `Value` into a standard numeric.

```{r warning=FALSE, message=FALSE}
fifa19$Value <- substr(fifa19$Value,2,200)
fifa19$ValueNum <- sapply(as.character(fifa19$Value), function(x) {
  unit <- substr(x, nchar(x), nchar(x))
  if (unit == "M") return (as.numeric(substr(x, 1, nchar(x)-1)) * 1000000)
  if (unit == "K") return (as.numeric(substr(x, 1, nchar(x)-1)) * 1000)
  as.numeric(x)
})

rownames(fifa19) <- make.names(fifa19$Name, unique = TRUE)
```

# Feature selection

Let's select only features related to player characteristics.

```{r warning=FALSE, message=FALSE}
fifa19_selected <- fifa19[,c(4,8,14:18,55:88,90)]
fifa19_selected$`Preferred Foot` <- factor(fifa19_selected$`Preferred Foot`)
```

# Feature engineering

Value is skewed. Will be much easier to model sqrt(Value).

```{r warning=FALSE, message=FALSE}
fifa19_selected$ValueNum <- sqrt(fifa19_selected$ValueNum)

fifa19_selected <- na.omit(fifa19_selected)
colnames(fifa19_selected) <- make.names(colnames(fifa19_selected))
```

# Create a gbm model

Let's use `gbm` library to create a `gbm` model with 250 trees 3 levels deep.

```{r warning=FALSE, message=FALSE}
set.seed(1313)

library("gbm")
fifa_gbm <- gbm(ValueNum~.-Overall, data = fifa19_selected, n.trees = 250, interaction.depth = 3)
```

# Create a DALEX explainer

Let's wrap gbm model into a DALEX explainer.

```{r warning=FALSE, message=FALSE}
library("DALEX")

fifa_gbm_exp <- explain(fifa_gbm, 
                        data = fifa19_selected, 
                        y = fifa19_selected$ValueNum^2, 
                        predict_function = function(m,x) 
                          predict(m, x, n.trees = 250)^2)
```

# Feature Importance explainer

Calculate Feature Importnace explainer.

```{r warning=FALSE, message=FALSE}
library("ingredients")
fifa_feat <- ingredients::feature_importance(fifa_gbm_exp)
plot(fifa_feat, max_vars = 12)
```

# Partial Dependency explainer

Calculate Partial Dependency explainer.

```{r warning=FALSE, message=FALSE}
fifa19_pd <- ingredients::partial_dependency(fifa_gbm_exp, variables = "Age")
plot(fifa19_pd)
```

# Ceteris Paribus explainer

Calculate Ceteris Paribus explainer.

```{r warning=FALSE, message=FALSE}
fifa19_cp_pg <- ingredients::ceteris_paribus(fifa_gbm_exp, new_observation = fifa19_selected["P..Gulácsi",], variables = "Age", variable_splits = list(Age = seq(18,45,0.1)))
plot(fifa19_cp_pg)
```

# Break Down explainer

Calculate Break Down explainer.

```{r warning=FALSE, message=FALSE}
library("iBreakDown")
fifa_pg_gbm <- break_down(fifa_gbm_exp, new_observation = fifa19_selected["P..Gulácsi",])
plot(fifa_pg_gbm)

fifa_pg_gbm$label = "Break Down for Péter Gulácsi (GBM model)"

library("ggplot2")
library("scales")
plot(fifa_pg_gbm, digits = 0) +  
  scale_y_continuous(labels = dollar_format(suffix = "€", prefix = ""), name = "Estimated value", limits = 1000000*c(2,13), breaks = 1000000*c(2.5,5,7.5,10))
```

# modelStudio app

Calculate modelStudio dashboard.

```{r eval = FALSE}
library("modelStudio")
options(
    parallelMap.default.mode        = "socket",
    parallelMap.default.cpus        = 4,
    parallelMap.default.show.info   = FALSE
)

fifa19_ms <- modelStudio(fifa_gbm_exp, 
                         new_observation = fifa19_selected[c("Cristiano.Ronaldo", "L..Messi","R..Lewandowski", "W..Szczęsny", "P..Gulácsi","A..Szalai", "Neymar.Jr"), ],
                         B = 5, N = 300,
                         parallel = TRUE,
                         options = modelStudioOptions(margin_left = 165),
                         digits = 0)

fifa19_ms

r2d3::save_d3_html(fifa19_ms, file = "fifa19_ms.html")
```

